---
title: "R Code Sample"
subtitle: "Senior Thesis Data Cleaning"
author: "Eric Huang"
date: "11/24/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(tidyr)
library(lubridate)
library(reshape2)
library(naniar)
library(gridExtra)
library(plyr)
library(knitr)
options(knitr.table.format = "html")
library(kableExtra)
```

## Overview

Analyzed University housing room assignment algorithm through from a market design perspective and presented findings at Carnegie Mellon’s Meeting of the Minds (annual research symposium) as well as Housing Services at an all-staff meeting in May 2019

### Raw Data

Source: Carnegie Mellon University Housing Services

+ 2017-18 Housing Room Selection Capacity/Occupancy numbers
    + Daily (almost daily) by 25+ residence halls (approx. 3500 bed spaces) 
+ 2018-19 Housing Room Selection Capacity/Occupancy numbers
    + Daily (almost daily) by 25+ residence halls (approx. 3500 bed spaces)
+ 2018 Housing Room Selection survey

## Summary Statistics

### Capacity/Occupancy

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# buildings total capacity
sum2 <- read.csv("~/Desktop/thesisDocuments/summary2.csv")
sum2 <- sum2[-c(27),-c(3:8)]

sum2 %>%
  kable() %>%
  kable_styling(full_width = F) %>%
  footnote(general = "2018: Total capacity of 3585")
```

### Eligible vs Actual Participants of Room Selection
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# capacity, eligibles, participation summary
sum1 <- read.csv("~/Desktop/thesisDocuments/summary1.csv")
colnames(sum1)[1] <- "variables"

customRows <- c("Total Capacity", "Capacity for RS", "Eligible for RS", "Eligible first years",
                "Eligible sophomores", "Eligible juniors", "Eligible seniors", "Participated in RS",
                "Participated first years", "Participated sophomores", "Participated juniors", 
                "Participated seniors")

sum1 %>%
  slice(match(customRows, variables)) %>%
  kable() %>%
  kable_styling(full_width = F) %>%
  add_indent(c(4:7,9:12)) %>%
  footnote(general = "NAs representing missing data, data that was either uncollected or unable to get from Housing.")
```

### Room Selection participation numbers by Phase

```{r, echo=FALSE, warning=FALSE, message=FALSE}
sum4_t <- read.csv("~/Desktop/thesisDocuments/summary4_t.csv")

colnames(sum4_t) <- c("Year", "Retention", "Block Housing", "General Selection", "Open Assignment", "Total")

years <- c("2014", "2015", "2016", "2017", "2018")

sum4_t %>%
  slice(match(years, Year)) %>%
  kable() %>%
  kable_styling(full_width = F)


```

## Code

```{r}  
# 2017-18 data

# Loading 2017-18 capacity/occupancy sheets
library("readxl")

sheets <- excel_sheets("~/Desktop/thesisDocuments/1718cap_occup.xlsx")
sheetList <- lapply(sheets, 
                    function(x)read_excel("~/Desktop/thesisDocuments/1718cap_occup.xlsx",
                                          range = "A1:P30",
                                          sheet = x))

names(sheetList) <- sheets

# Selecting upperclass houses (rows)
uc_houses_num <- c(2,3,5,7,8,9,11,13,15,16,18,20,21,23,24,25,26,27,29)
sheetNames <- names(sheetList)
uc_list <- list()
for (i in 1:length(sheetNames)) {
  curr_sheet <- sheetList[[sheetNames[i]]]
  if (curr_sheet[21,1] == "Stever House"){
    uc_houses_num <- c(2,3,5,7,8,9,11,13,15,16,18,20,22,23,24,25,26,27)
  } else {
    uc_houses_num <- c(2,3,5,7,8,9,11,13,15,16,18,20,21,23,24,25,26,27)
  }
  
  cleaned_sheet <- curr_sheet[uc_houses_num,]
  uc_list[[i]] = cleaned_sheet
}
names(uc_list) <- sheets
list2env(uc_list,.GlobalEnv)

# Selecting upperclass occupancy/vacancy variables (columns)
sheetNames <- names(uc_list)
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list[[sheetNames[i]]]
  cleaned_sheet <- curr_sheet[c(1,4,8,12)]
  uc_list[[i]] = cleaned_sheet
}

names(uc_list) <- sheets
list2env(uc_list,.GlobalEnv)

# Computing column totals
sheetNames <- names(uc_list)
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list[[sheetNames[i]]]
  cleaned_sheet <- rbind(curr_sheet, c("Total", colSums(curr_sheet[-1])))
  colnames(cleaned_sheet) <- c("Building", "Capacity (in beds)", "Actual Occupancy", "Vacancy")
  uc_list[[i]] = cleaned_sheet
}

names(uc_list) <- sheets
list2env(uc_list,.GlobalEnv)

# Editing row name wording
sheetNames <- names(uc_list)
# may need to run loop twice for *both* 'Spirit' *and* '5170 MM' to come up
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list[[sheetNames[i]]]
  cleaned_sheet <- curr_sheet
  if (curr_sheet[12,1] == "Spirit House/Roselawn 15"){
    cleaned_sheet[12,1] = "Spirit House"
  } else if (curr_sheet[13,1] == "Spirit House/5170 MM"){
    cleaned_sheet[13,1] = "5170 MM"
  } 
  uc_list[[i]] = cleaned_sheet
}

names(uc_list) <- sheets
list2env(uc_list,.GlobalEnv)

# Adjusting 'Building' order
sheetNames <- names(uc_list)
house_order <- c("Clyde House",
                 "Doherty Apartments",
                 "Fairfax Apartments",
                 "Henderson House",
                 "The Highlands Apts.",
                 "Margaret Morrison Apts",
                 "McGill House",
                 "Morewood Gardens",
                 "Neville Apartments",
                 "Resnik House",
                 "Roselawn Houses",
                 "Shady Oak Apartments",
                 "Spirit House",
                 "Webster Apartments",
                 "Welch House",
                 "West Wing",
                 "Woodlawn Apartments",
                 "5170 MM",
                 "Total")

for (i in 1:length(sheetNames)){
  sheet <- uc_list[[sheetNames[i]]]
  sheet <- sheet[match(house_order, sheet$Building),]
  uc_list[[i]] = sheet 
}

names(uc_list)
list2env(uc_list,.GlobalEnv)

# removing NA row from 'Original - 10-26-16' sheet
uc_list$`Original - 10-26-16` <- uc_list$`Original - 10-26-16`[c(1:17,19),]
names(uc_list)
list2env(uc_list,.GlobalEnv)

# reodering 'Original - 10-26-16' from position 1 to position 62
uc_list <- c(uc_list[-1],uc_list[1])
names(uc_list)
list2env(uc_list,.GlobalEnv)

# removing inaccurate "total" row from 'Original - 10-26-16' sheet
uc_list[[62]] <- uc_list[[62]][-c(18),]
curr_sheet <- uc_list[[62]]
names(uc_list)
list2env(uc_list,.GlobalEnv)

# recalculating "total" for 'Original - 10-26-16' sheet
curr_sheet <- uc_list[[62]]
curr_sheet2 = curr_sheet[[1]]
curr_sheet2 = as.data.frame(matrix(nrow=length(curr_sheet2), ncol=4))
curr_sheet2$Building = curr_sheet[[1]]
curr_sheet2$Capacity = as.numeric(curr_sheet[[2]])
curr_sheet2$Occ = as.numeric(curr_sheet[[3]])
curr_sheet2$Vac = as.numeric(curr_sheet[[4]])
  
curr_sheet2 = curr_sheet2[,-(1:4)]
colnames(curr_sheet2) = colnames(curr_sheet)
cleaned_sheet <- rbind(curr_sheet, c("Total", colSums(curr_sheet2[-1])))
uc_list[[62]] = cleaned_sheet

names(uc_list)
list2env(uc_list,.GlobalEnv)

# reorder dataframes by chronological order
library(rlist)
uc_list <- list.reverse(uc_list)
names(uc_list)
list2env(uc_list,.GlobalEnv)

```

```{r}
# 2018-19 data

# Loading 2018-19 capacity/occupancy sheets
sheets1 <- excel_sheets("~/Desktop/thesisDocuments/1819cap_occup.xlsx")
sheetList1 <- lapply(sheets1, 
                    function(x)read_excel("~/Desktop/thesisDocuments/1819cap_occup.xlsx",
                                          range = "A1:P30",
                                          sheet = x))
names(sheetList1) <- sheets1

# Selecting upperclass houses (rows)
sheetNames <- names(sheetList1)
uc_list1 <- list()
for (i in 1:length(sheetNames)) {
  curr_sheet <- sheetList1[[sheetNames[i]]]
  if (curr_sheet[17,1] == "Scobell House"){
    uc_houses_num <- c(2,3,5,7,8,9,11,13,15,16,18,20,21,23,24,25,26,27)
  } else {
    uc_houses_num <- c(2,3,5,7,8,9,10,12,14,16,17,18,20,22,24,25,26,27)
  }
  cleaned_sheet <- curr_sheet[uc_houses_num,]
  uc_list1[[i]] = cleaned_sheet
}

names(uc_list1) <- sheets1
list2env(uc_list1,.GlobalEnv)

# Selecting upperclass occupancy/vacancy variables (columns)
sheetNames <- names(uc_list1)
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list1[[sheetNames[i]]]
  cleaned_sheet <- curr_sheet[c(1,4,8,12)]
  uc_list1[[i]] = cleaned_sheet
}

names(uc_list1) <- sheets1
list2env(uc_list1,.GlobalEnv)

# Computing column totals
sheetNames <- names(uc_list1)
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list1[[sheetNames[i]]]
  
  curr_sheet2 = curr_sheet[[1]]
  curr_sheet2 = as.data.frame(matrix(nrow=length(curr_sheet2), ncol=4))
  curr_sheet2$Building = curr_sheet[[1]]
  curr_sheet2$Capacity = as.numeric(curr_sheet[[2]])
  curr_sheet2$Occ = as.numeric(curr_sheet[[3]])
  curr_sheet2$Vac = as.numeric(curr_sheet[[4]])
  
  curr_sheet2 = curr_sheet2[,-(1:4)]
  colnames(curr_sheet2) = colnames(curr_sheet)
  
  cleaned_sheet <- rbind(curr_sheet, c("Total", colSums(curr_sheet2[-1])))
  colnames(cleaned_sheet) <- c("Building", "Capacity (in beds)", "Actual Occupancy", "Vacancy")
  uc_list1[[i]] = cleaned_sheet
}

names(uc_list1) <- sheets1
list2env(uc_list1,.GlobalEnv)

# Editing row name wording
sheetNames <- names(uc_list1)
# may need to run loop twice for *both* 'Spirit' *and* '5170 MM' to come up
for (i in 1:length(sheetNames)){
  curr_sheet <- uc_list1[[sheetNames[i]]]
  cleaned_sheet <- curr_sheet
  if (curr_sheet[11,1] == "Roselawn 15/Old Spirit"){
    cleaned_sheet[11,1] = "Spirit House"
  } else if (curr_sheet[12,1] == "Spirit House/Roselawn 15"){
    cleaned_sheet[12,1] = "Spirit House"
  } else if (curr_sheet[14,1] == "Spirit House/5170 MM"){
    cleaned_sheet[14,1] = "5170 MM"
  } else if (curr_sheet[13,1] == "Spirit House/5170 MM"){
    cleaned_sheet[13,1] = "5170 MM"
  }
  uc_list1[[i]] = cleaned_sheet
}

names(uc_list1) <- sheets1
list2env(uc_list1,.GlobalEnv)

# Adjusting 'Building' order
sheetNames <- names(uc_list1)
house_order <- c("Clyde House",
                 "Doherty Apartments",
                 "Fairfax Apartments",
                 "Henderson House",
                 "The Highlands Apts.",
                 "Margaret Morrison Apts",
                 "McGill House",
                 "Morewood Gardens",
                 "Neville Apartments",
                 "Resnik House",
                 "Roselawn Houses",
                 "Shady Oak Apartments",
                 "Spirit House",
                 "Webster Apartments",
                 "Welch House",
                 "West Wing",
                 "Woodlawn Apartments",
                 "5170 MM",
                 "Total")

for (i in 1:length(sheetNames)){
  sheet <- uc_list1[[sheetNames[i]]]
  sheet <- sheet[match(house_order, sheet$Building),]
  uc_list1[[i]] = sheet 
}

names(uc_list1)
list2env(uc_list1,.GlobalEnv)

# removing '2-21-18(day1RS)' and '2-22-18(day2RS)' sheets from general list
uc_list1 <- uc_list1[-c(51,52)]
names(uc_list1)
list2env(uc_list1,.GlobalEnv)

# create separate list for the "proposed" sheets
proposed_list <- uc_list1[c(50:55)]
names(proposed_list)
list2env(proposed_list,.GlobalEnv)

# removing "proposed" sheets from general list and reordering dataframes by chronological order
uc_list1 <- uc_list1[-c(50:55)]
uc_list1 <- list.reverse(uc_list1)
names(uc_list1)
list2env(uc_list1,.GlobalEnv)

```